Continuous developmental change can explain discontinuities in word learning

### 2019-07-22 journal/ms.tex
journal v1

\abstract{
``Cognitive development is often characterized in term of
discontinuities, but these discontinuities can sometimes be apparent
rather than actual and can arise from continuous developmental change.
To explore this idea, we use as a case study the finding by Stager and
Werker (1997) that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. Early
explanations proposed that children may not be able to encode subtle
phonetic contrasts when learning novel word meanings, thus suggesting a
discontinuous/stage-like pattern of development. However, later work has
revealed (e.g., through using simpler testing methods) that children do
encode such contrasts, thus favoring a continuous pattern of
development. Here we propose a probabilistic model describing how
development may proceed in a continuous fashion across the lifespan. The
model accounts for previously documented facts and provides new
predictions. We collected data from preschool children and adults, and
we showed that the model can explain various patterns of learning both
within the same age and across development. The findings suggest that
major aspects of cognitive development that are typically thought of as
discontinuities, may emerge from simpler, continuous mechanisms.''

@@

We use a case study from word learning literature. Stager and Werker
(1997) first showed that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills.
Indeed, though infants around 14-month old can distinguish similar sound
pairs such as \enquote{dih} and \enquote{bih}, they appear to fail in
mapping this pair to two different objects.

By around 17 m.o, children succeed in the same task (Werker, Fennell,
Corcoran, \& Stager, 2002). How does development proceed? The answer to
this question depends on the nature of the chidlren's word encoding in
the task. Early accounts assumed that children encode words in a rather
binary way: they either fail or succeed in attending to the relevant
phonetic details (simultaneously with the meanings). In the orginal
paper, Stager and Werker (1997) noted that \enquote{Infants of 14 months
fail to detect the same phonetic detail that they can easily detect in a
simple syllable discrimination task}. They argued that the high cogntive
demands of the task make children experience a limitation in the amount
of phonetic details they can access. In a similar vein, Pater et al.
(2004) argued that \enquote{when a contrast is first acquired, it is not
stable and can be lost under processing demands}. This binary account
suggested a discontinuous/stage-like pattern of development whereby
younger children fail to encode the contrastive phonetic detail, whereas
older children succeed.

Subsequent findings have investigated the implications of this
hypothesis in both infants and adults. First, if the demands of the task
totally empede the encoding of subtle phonetic contrasts, then infants
should fail in learning the mappings between the sounds and the objects,
regardless of how this learning is probed. However, 14-month-olds
succeed in this task under the same \emph{learning} conditions as in
Stager and Werker (1997), but when an easier \emph{testing} method is
used (Yoshida, Fennell, Swingley, \& Werker, 2009). Second, if the
minsmatch between sound distrimination and word learning is only related
to limited cognitive resources in infancy, then adults --- who have
greater attentianal and working memory capacity --- should always
succeed in mapping pairs of sounds they can discriminate to different
objects. Nonetheless, even adults show patterns learning that mirror
those shown by 14-month-olds (Pajak, Creel, \& Levy, 2016; White, Yee,
Blumstein, \& Morgan, 2013).

This new set of evidence points towards another scenario, where the
representations are encoded in a probabilistic (rather than binary) way,
and where development is continuous, rather than stage-like (see also
Swingley, 2007). On this account, correct representations are learned
early in development, but these representations are encoded with higher
uncertainty in younger children, leading to \emph{apparent} failure in
relatively demanding tasks. Development is a continuous process whereby
the initial noisy representations become more precise. Crucially, in a
probabilsitic account, more precise representation do not mean they
perfect, thus accounting for the fact that even adults show low accuracy
learning when the sounds are subtle, e.g., non-native (Pajak et al.,
2016).

We provide an intuitive illustration of how such an account explains
patterns of learning and development in Figure \ref{fig:illus}. We
observe low accuracy in word learning when the perceptual distance
between the labels is small relative to the uncertainty with which these
labels are encoded. For example, in Stager and Werker's original
experiment, children are supposed to associate label 1 (\enquote{bih})
and label 2 (\enquote{dih}) with object 1 and object 2, respectively.
Though infants could learn that the label \enquote{bih} is a better
match to object 1 than \enquote{dih}, they could still judge the sound
\enquote{dih} as a plausible instance of the label \enquote{bih}, thanks
to the relatively large uncertainty of the encoding, and this confusion
leads to \enquote{failure} in the recognition task. According to this
account, accuracy in word learning improves if we increase either the
perceptual distinctiveness of the stimuli (e.g., through using
different-sounding labels), or the precision of the encoding itself
(e.g., across development).

Building on this intuition, the current work proposes a probabilistic
model, which we use to both account for previous experimental findings,
and to make new predictions that have not been tested before. Using new
data collected from both preschool children and adults, we show that the
model can explain various patterns of learning both within the same age
and across development.

@@

The learner hears a sound \(s\) and has to decide which object \(o\)
provides an optimal match to this sound. To this end, they must compute
the probability \(P(o|s)\) for all possible objects. This probability
can be computed by summing over all possible concepts and labels:
\[P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) \]

@@

\begin{equation}
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\end{equation}

@@

We use the model to predict performance in the word learning task
introduced by Stager and Werker (1997), with a two-alternative forced
choice as in Yoshida et al. (2009). In this task, participants are first
exposed to the association between pairs of nonsense words (e.g.,
\enquote{lif}/\enquote{neem}) and pairs of objects. The word-object
associations are introduced sequentially. After this exposure phase,
participants perform a series of test trials. In each of these trials,
one of the two sounds is uttered (e.g., \enquote{lif}) and participants
choose the corresponding object from the two alternatives. An overview
of the task is shown in Figure \ref{fig:task}.

We used Equation 1 and the probability distributions defined above to
obtain the exact analytical expression for the probability of accurate
responses \(p(o_T | s)\) (target object \(o_T\) given a sound \(s\)) in
the simple case of two-alternative forced choice in the testing phase of
our experimental task:

\begin{equation}
P(o_T|s)= \frac{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}}{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}+ e^{-\Delta s^2 /2\sigma_L^2} + e^{-\Delta o^2 /2\sigma_C^2 }}
\end{equation}

Figure \ref{fig:simulation} show simulations of the predicted accuracy
(Expression 2) as a function of the distinctiveness parameters
(\(\Delta s\) and \(\Delta o\)) and the precision parameters, i.e., the
variances of the distributions \(p(s| L)\) and \(p(o | C)\). To
understand the qualitative behavior of the model, we assumed for
simplicity that the precision parameter has similar values in both
distributions, i.e., \(\sigma =\sigma_C \approx \sigma_L\) (but we will
allow those parameters to vary independently in the rest of the paper).

@@

The simulations explain some previously documented facts, and make new
predictions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty (characterized with \(\sigma\))
  decreases. This fact may explain development, i.e., younger children
  have noisier representations (see Swingley, 2007; Yoshida et al.,
  2009), which leads to lower word recognition accuracy, especially for
  similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (e.g., Fennell \& Waxman, 2010;
  Rost \& McMurray, 2009; Thiessen, 2007), but to our knowledge no
  previous study in the literature tested the effect of the visual
  distance between the semantic referents.
\end{enumerate}

@@

In this experiment, we tested participants in the word learning task
introduced above (Figure \ref{fig:task}). More precisely, we explored
the predictions related to both distinctiveness and precision. Sound
similarity (\(\Delta s\)) and object similarity (\(\Delta o\)) were
varied simultaneously in a within-subject design. Two age groups
(preschool children and adults) were tested on the same task to explore
whether development can be characterized with the uncertainty
parameters, \(\sigma_C\) and \(\sigma_L\). The experiment, sample size,
exclusion criteria and the model's main predictions were pre-registered.

@@

We planned to recruit a sample of N=60 children ages 4-5 years from the
Bing Nursery School on Stanford University's campus. Here we report data
from n=55 children. An additional n=35 children participated but were
removed from analyses because they were not above chance on the catch
trials due to the challenging nature of our procedure (see below). We
also planned to recruit a sample of N=60 adults on Amazon Mechanical
Turk. Data from n=26 participants were excluded due to low scores on the
catch trials (n=26 ) or because they were familiar with the non-English
sound stimuli we used in the adult experiment (n=0), yielding a final
sample of n=74.

@@

The sound stimuli were generated using the MBROLA Speech Synthesizer
(Dutoit, Pagel, Pierret, Bataille, \& Van der Vrecken, 1996). We
generated three kinds of nonsense word pairs which varied in their
degree of similarity to English speakers: 1) \enquote{different}:
\enquote{lif}/\enquote{neem} and \enquote{zem}/\enquote{doof}, 2)
\enquote{intermediate}: \enquote{aka}/\enquote{ama} and
\enquote{ada}/\enquote{aba}, and 3) \enquote{similar} non-English
minimal pairs: \enquote{ada}/\enquote{a\textipa{d\super h}a} (in hindi)
and \enquote{a\textipa{Q}a}/\enquote{a\textipa{\textcrh}a} (in arabic).

As for the objects, we used the Dynamic Stimuli javascript
library\footnote{https://github.com/erindb/stimuli} which allowed us to
generate objects in four different categories: \enquote{tree},
\enquote{bird}, \enquote{bug}, and \enquote{fish}. These categories are
supposed to be naturally occurring kinds that might be seen on an alien
planet. In each category, we generated \enquote{different},
\enquote{intermediate} and \enquote{similar} pairs by manipulating a
continuous property controlling features of the category's shape (e.g,
body stretch or head fatness).

In a separate survey, \(N=20\) participants recruited on Amazon
Mechanical Turk evaluated the similarity of each sound and object pair
on a 7-point scale. We scaled responses within the range {[}0,1{]}. Data
are shown in Figure \ref{fig:stim}, for each stimulus group. These data
will be used in the models as the perceptual distance of sound pairs
(\(\Delta s\)) and object pairs (\(\Delta o\)).

\includegraphics{ms_files/figure-latex/unnamed-chunk-6-1.pdf}
\#\#\#Design

Each age group saw only two of the three levels of similarity described
in the previous sub-section: \enquote{different} vs.
\enquote{intermediate} for preschoolers, and \enquote{intermediate} vs.
\enquote{similar} for adults. We made this choice in light of pilot
studies showing that adults were at ceiling with \enquote{different}
sounds/objects, and children were at chance with the \enquote{similar}
sounds/objects. That said, this difference in the level of similarity is
accounted for in the model through using the appropriate perceptual
distance used in each age group (Figure \ref{fig:stim}).

@@

\caption{Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both experimental results (solid lines) and model predictions (dashed lines). Error bars represent 95\% confidence intervals.}\label{fig:allData}
\end{figure}

Experimental results are shown in Figure \ref{fig:allData} (solid
lines). We first analyzed the results using a mixed-effects logistic
regression with sound distance, object distance and age group as fixed
effects, and with a maximal random effects structure (allowing us to
take into account the full nested structure of our data) (Barr, Levy,
Scheepers, \& Tily, 2013). We found main effects for all the fixed
effects in the regression. For the sound distance, we obtained
\(\beta =\) 0.52 (\(p\) \textless{} 0.001), replicating previous
findings. For object distance, we found \(\beta =\) 0.83 (\(p\)
\textless{} 0.001), and this finding confirms the new prediction of our
model. Finally, for the age group, we obtained \(\beta =\) 0.76 (\(p\)
\textless{} 0.001), showing that performance improves with age. In
addition, we found two-way interactions between sound distance and age
(\(\beta =\) 0.15, \(p\) = 0.17) and between object distance and age
(\(\beta =\) 0.45, \(p\) \textless{} 0.001).

We next fit our model (using Equation 2) to the participants' responses
in each age group. The values of \(\Delta s\) and \(\Delta o\) were set
based on data from the similarity judgment task (Figure \ref{fig:stim}).
Thus, the model has two degrees of freedom for each group, i.e.,
\(\sigma_C\) and \(\sigma_L\). Figure \ref{fig:allData} (dashed lines)
shows the predictions. The model captures the qualitative patterns in
both age groups: starting from a low accuracy recognition when both the
sound and object distances are small, the model correctly predicts an
increase in accuracy when either the sound distance or the object
distance increases. Further, accuracy is correctly predicted to be
maximal when both the sound and object distances are high.

The values of the parameters were as follows. Children had a
label-specific uncertainty of \(\sigma_S =\) 0.83 {[}0.64,
1.02{]}\footnote{All uncertainty intervals in this paper represent 95\% Confidence Intervals.},
and a concept-specific uncertainty of \(\sigma_C =\) 0.31 {[}0.11,
0.51{]}. Adults had a label-specific uncertainty of \(\sigma_S =\) 0.12
{[}0.12, 0.13{]}, and a concept-specific uncertainty of \(\sigma_C =\)
0.17 {[}0.16, 0.18{]}. As predicted, the uncertainty parameters were
larger for children than they were for adults, showing that the
probabilistic representations becomes more refined (that is, \(\sigma\)
becomes smaller) across development. The developmental effect was more
important for the label-specific uncertainty.

The models explained the majority of the variance in the participants'
mean responses (\(R^2=\) 1, for the combined adult and children's data).
To investigate whether the model's predictive power was due to
overfitting, we fit a simplified version with only one degree of freedom
(i.e., a single variance common to both sounds and objects). This
single-variance model explained as much variance in the mean responses
(\(R^2=\) 0.95). It also captured the main qualitative patterns (graph
not shown), suggesting that the explanatory power of the model is
largely due to its structure, rather than its degrees of freedom.

An unexpected outcome was that adult participants deviated slightly from
the model's numbers: While the model predicted accuracy to be more
sensitive to object distance when sound distance is higher (and
vice-versa), adult participants showed the opposite pattern. This
deviation is an interesting starting point for future work as it may
suggest that participants are more likely to pay attention to and
integrate additional sources of information when ambiguity is higher.

@@

This paper explored the idea that some seemingly stage-like patterns in
cognitive development can be characterized in a continuous fashion. We
used as a case study the seminal work of Stager and Werker (1997)
showing a discrepancy between children's speech perception abilities and
their word learning skills. While much of the previous investigation of
this finding has been interested in the source of this discrepancy, here
we have explored how it could arise from continuous developmental change
in perceptual uncertainty.

Building on some previous discussions (e.g., Swingley, 2007; Yoshida et
al., 2009), we proposed a model where perceptual stimuli are encoded
probabilistically. We tested the model's predictions against data
collected from preschool children and adults and we showed that
developmental changes in word-object mappings can indeed be
characterized as a continuous refinement (i.e., uncertainty reduction)
in qualitatively similar representations across the life span.

>>>>

\abstract{
``Cognitive development is often characterized in term of
discontinuities, but these discontinuities can sometimes be apparent
rather than actual and can arise from continuous developmental change.
To explore this idea, we use as a case study the finding by Stager and
Werker (1997) that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. Early
explanations proposed that children may not be able to encode subtle
phonetic contrasts when learning novel word meanings, thus suggesting a
discontinuous/stage-like pattern of development. However, later work has
revealed (e.g., through using simpler testing methods) that children do
encode such contrasts, thus favoring a continuous pattern of
development. Here we propose a probabilistic model describing how
development may proceed in a continuous fashion across the lifespan. The
model accounts for previously documented facts and provides new
predictions. We collected data from preschool children and adults, and
we showed that the model can explain various patterns of learning both
within the same age and across development. The findings suggest that
major aspects of cognitive development that are typically thought of as
discontinuities may emerge from simpler, continuous mechanisms.''

@@

We use a case study from word learning literature. Stager and Werker
(1997) first showed that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. The
authors measured word learning using an audio-visual habituation Switch
task. First, infants are familiarized with two word-object pairings
(e.g., Word A with Object A and Word B with Object B). Second, they are
tested using two types of trials. The control \enquote{same} trial
consists of a correct pairing (e.g., Word A with Object A) and the
\enquote{switch} trial consists of a wrong pairing (e.g., Word A with
Object B). If babies have correctly learned the association during the
familiarization, they are supposed to be surprised by the
\enquote{switch} trial and not by the \enquote{same} trial. The former
should thus result in a greater looking time compared to the latter
(Werker, Cohen, Lloyd, Casasola, \& Stager, 1998).

Though infants around 14-month old can distinguish perceptually similar
sound pairs such as \enquote{dih} and \enquote{bih}, they appear to fail
in mapping this pair to two different objects in the switch task. This
failure has initially been taken as evidence that 14-month olds do not
encode subtle sounds during meaning learning (Pater, Stager, \& Werker,
2004; Stager \& Werker, 1997). This interpretation suggested a
discontinuous/stage-like pattern of development whereby younger children
fail to encode the contrastive phonetic detail, whereas older children,
around 17 months (Werker, Fennell, Corcoran, \& Stager, 2002), typically
do.

The initial discontinuous interpretation has been challenged by
subsequent work. For instance, Yoshida, Fennell, Swingley, and Werker
(2009) investigated whether failure in the switch task reflects a lack
of sound encoding during \emph{familiarization}, or whether it is only
due to the demands of the \emph{testing} method which does not allow
learning below a certain threshold to be detected. They used the same
familiarization procedure as Stager and Werker (1997), but instead of
comparing the looking times in \enquote{same} and \enquote{switch}
trials, they tested infants using a two-alternative choice task
comparing fixations to target and distractor objects (Fernald, Perfors,
\& Marchman, 2006; Golinkoff, Hirsh-Pasek, Cauley, \& Gordon, 1987).
Using this simpler and finer-grained task, the researcher found evidence
for learning even in 14-month olds.

Another challenge to the discontinuous account of development came from
adult studies. If the mismatch between sound discrimination and word
learning is only a stage in early infancy, then this mismatch should
disappear by adulthood. Nonetheless, even adults show patterns of
learning that mirror those shown by 14-month-olds when the sound
contrasts more challenging (Pajak, Creel, \& Levy, 2016; White, Yee,
Blumstein, \& Morgan, 2013).

Some researchers (Pajak et al., 2016; e.g., Swingley, 2007; Yoshida et
al., 2009) proposed that the phonological form may not be encoded in a
binary fashion, i.e., it is not the case that children either succeed or
fail in encoding minimal contrast when learning the meanings. Rather,
they may be encoding the phonological form of words in a probabilistic
fashion. According to this view, development does not so much involve a
qualitative shift (i.e., a sudden emergence of an ability that did not
exist before) as much as it consists in the continuous refinement of
initially noisy representations.

In a probabilistic account, a word can be represented as a probability
distribution over sound instances organized in a similarity space. The
probability is highest at the most typical sound instance. It decreases
as the instance becomes less typical. The precision of the
representation can be characterized by how much it tolerates slightly
atypical pronunciations. This tolerance is captured formally by the
variance of the probability distribution: larger variance indicates
higher tolerance and lower precision, whereas smaller variance indicates
lower tolerance and higher precision.

This framework can explain several findings. In Stager and Werker's
original experiment, children are supposed to associate one label
\enquote{bih} with object 1 and a second label \enquote{dih} with object
2. Though infants could learn that the label \enquote{bih} is a better
match to object 1 than \enquote{dih}, they could still judge the sound
\enquote{dih} as a plausible instance of the label \enquote{bih}, thanks
to the relatively large variance of the early encoding, and this
confusion leads to \enquote{failure} in the recognition task (Figure
\ref{fig:illus}, top). Though learning is small and is easily disrupted
by the Switch task, it can still be detected when less demanding methods
are used (Yoshida et al., 2009).

The learning accuracy increases (and is detected even by demanding tests
such as the Switch) for more distinct word-forms (e.g., \enquote{lif}
vs. \enquote{neem}) where the perceptual distance is large relative to
the variance (Figure \ref{fig:illus}, left). Distinctiveness can be
enhanced even for minimally different sounds when other cues highlight
their difference (Dautriche, Swingley, \& Christophe, 2015; Rost \&
McMurray, 2009, 2010; Thiessen, 2007; Yeung \& Werker, 2009).

Finally, development can be understood as an increase in the precision
(i.e., a decrease in the variance) of the probabilistic representations
(Figure \ref{fig:illus}, right). Such an increase in precision renders
minimally different sounds less confusing. Importantly, a more precise
representation still has a non-zero variance --- Learning difficulties
can still be induced with challenging stimuli or in cognitively
demanding situations as was demonstrate in adults studies (Pajak et al.,
2016; White et al., 2013).

\subsection{This study}\label{this-study}

The probabilistic account has been put forward to explain patterns of
learning and development at the qualitative level. However, it is
crucial to have a precise computational instantiation of this account
which would allow us to \emph{quantify} its explanatory power. We could
find one previous work that attempted to provide such a computational
instantiation (Hofer and Levy, 2017). However, this previous work was
designed with the goal of reproducing the results of a specific study
(Pajak et al., 2016) which focused on explaining the mismatch between
speech perception and word learning in adults rather than on exploring
the mechanism of development.

The present work proposes a model of word-pair learning based on the
probabilistic account. We tested the ability of this model to both
\emph{explain} various findings in previous experiments in both children
and adults (e.g., the fact that similar words are harder to learn than
different words) and to \emph{predict} new learning patterns that have
not been tested before (i.e., the effect of the referents' similarity on
word learning). Crucially, we explored the extent to which development
can be understood as a continuous refinement in similar representations
across the lifespan.

The paper is organized as follows. First, we introduce the model and we
explain how it allows us to characterize behavior in a word learning
task which resembles the one used in Stager and Werker (1997) and
Yoshida et al. (2009). Then we explore the predictions of the model
through simulating its behavior across different parameter settings.
Next we quantify the extent to which the model's predictions account for
human data we collected from both preschool children and adults.
Finally, we discuss the results in the lights of existing accounts of
word development.

@@

The learner hears a sound \(s\) and has to decide which object \(o\)
provides an optimal match to this sound (see Figure \ref{fig:task}). To
this end, they must compute the probability \(P(o|s)\) for all possible
objects. This probability can be computed by summing over all possible
concepts and labels:
\[P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) \]

@@

\[
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\]

@@

We use the model to predict word learning in a task similar to the one
introduced by Stager and Werker (1997). We used a modified version of
the task where the testing method consists in a two-alternative forced
choice (Yoshida et al., 2009). In this task, participants are first
exposed to the association between pairs of nonsense words (e.g.,
\enquote{lif}/\enquote{neem}) and pairs of objects. The word-object
associations are introduced sequentially. After this exposure phase,
participants perform a series of test trials. In each of these trials,
one of the two sounds is uttered (e.g., \enquote{lif}) and participants
choose the corresponding object from the two alternatives. An overview
of the task is shown in Figure \ref{fig:task}.

From the general expression (1), we derive three exact analytical
solutions instantiating different learning assumptions. The first
solution is derived by assuming that the labels are recovered from
sounds with a certain level of uncertainty \(\sigma_L\), but that
concepts are unambiguously recovered from the observed objects, i.e.,
\(\sigma_C \rightarrow 0\). This assumption has been made --- whether
implicitly or explicitly --- by most previous work in this line of
research. One important implication of this assumption is that only the
similarity of word sounds modulates success in word learning, not the
similarity of the referents (as long as these referents are
differentiated perceptually). This assumption yields the following
probability function:

\begin{equation}
P(o_T|s)= \frac{1}{1 + e^{-\frac{\Delta s^2}{2\sigma_L^2}}}
\end{equation}

The second solution is derived by making the more general assumption
that both the labels and the concepts are recovered with ambiguity from
the sounds and objects. We first introduce the simplifying assumption
that the label-related uncertainty \(\sigma_L\) and the concept-related
uncertainty \(\sigma_C\) are of a similar magnitude, i.e.,
\(\sigma_C \approx \sigma_L = \sigma\). This assumption makes the
prediction that the sound similarity and the object similarity impact
word learning accuracy in exactly the same way. Furthermore, it allows
us to study the behavior of the model with minimal free parameters.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- \frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}}{1 + e^{-\frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}+ e^{-\frac{\Delta s^2}{2\sigma^2}} + e^{-\frac{\Delta o^2}{2\sigma^2}}}
\end{equation}

We finally derive the third (and most general) solution which allows
label- and concept-related uncertainties to vary independently.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- (\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}}{1 + e^{-(\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}+ e^{-\frac{\Delta s^2}{2\sigma_L^2}} + e^{-\frac{\Delta o^2}{2\sigma_C^2}}}
\end{equation}

In order to understand the predictions of the models (especially the
more general ones, i.e., Model 2 and 3), Figure \ref{fig:simulation}
show some simulations of the accuracy \(P(o_T|s)\) as a function of the
distinctiveness parameters (\(\Delta s\) and \(\Delta o\)) and the
uncertainty parameters \(\sigma_L\) and \(\sigma_C\).

@@

The simulations explain two experimental results from previous studies
and make one new prediction:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty \(\sigma\) decreases. This fact
  may explain development, i.e., younger children have noisier
  representations (see Swingley, 2007; Yoshida et al., 2009), which
  leads to lower word recognition accuracy, especially for
  similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (e.g., Fennell \& Waxman, 2010;
  Rost \& McMurray, 2009; Thiessen, 2007), but to our knowledge, no
  previous study in the literature tested the effect of the visual
  distance between the semantic referents.
\end{enumerate}

@@

In this experiment, we tested participants in the word learning task
introduced above (Figure \ref{fig:task}). More precisely, we explored
the predictions related to both distinctiveness and precision. Sound
similarity (\(\Delta s\)) and object similarity (\(\Delta o\)) were
varied simultaneously in a within-subject design. Two age groups
(preschool children and adults) were tested on the same task to explore
whether development can be characterized with the uncertainty
parameters, \(\sigma_C\) and \(\sigma_L\). The experiment, sample size,
exclusion criteria and the model's main predictions were
pre-registered.\footnote{https://osf.io/942gv/}

@@

We report data from \(N=\) 63 children ages 4-5 years from the Bing
Nursery School on Stanford University's campus. An additional \(N=\) 39
children participated but were removed from analyses because they were
not above chance on the catch trials due to the challenging nature of
our procedure (see below). We also report data from \(N=\) 74 adult
participants tested on Amazon Mechanical Turk. An additional \(N=\) 26
were tested but removed from analyses because they had low scores on the
catch trials or because they were familiar with the non-English sound
stimuli we used in the adult experiment.

@@

The sound stimuli were generated using the MBROLA Speech Synthesizer
(Dutoit, Pagel, Pierret, Bataille, \& Van der Vrecken, 1996). We
generated three kinds of nonsense word pairs which varied in their
degree of perceptual similarity to English speakers: 1) \emph{different}
pairs: \enquote{lif}/\enquote{neem} and \enquote{zem}/\enquote{doof}, 2)
\emph{intermediate} pairs: \enquote{aka}/\enquote{ama} and
\enquote{ada}/\enquote{aba}, and 3) \emph{similar} non-English pairs:
\enquote{ada}/\enquote{a\textipa{d\super h}a} (in hindi) and
\enquote{a\textipa{Q}a}/\enquote{a\textipa{\textcrh}a} (in arabic).

As for the objects, we used the Dynamic Stimuli javascript
library\footnote{https://github.com/erindb/stimuli} which allowed us to
generate objects in four different categories: \enquote{tree,}
\enquote{bird,} \enquote{bug,} and \enquote{fish.} These categories are
supposed to be naturally occurring kinds that might be seen on an alien
planet. In each category, we generated \emph{different},
\emph{intermediate} and \emph{similar} pairs by manipulating a
continuous property controlling features of the category's shape (e.g,
body stretch or head fatness).

In order to validate and quantify our similarity scales, we ran a
separate survey on Amazon Mechanical Turk where we asked \(N=20\) adults
participants to evaluate the similarity of each sound and object pair on
a 7-point scale. Data are shown in Figure \ref{fig:stim} where we scaled
responses within the range {[}0,1{]} for each stimulus group. These data
will be used in all models as a proxy for the perceptual distance
between the sound pairs (\(\Delta s\)) and the object pairs
(\(\Delta o\)).

\subsubsection{Design}\label{design}

Each age group saw only two of the three levels of similarity described
in the previous sub-section: \emph{different} vs. \emph{intermediate}
for the preschoolers, and \emph{intermediate} vs. \emph{similar} for
adults. We made this choice in light of pilot studies showing that
adults were at ceiling with \emph{different} sounds/objects, and
children were at chance with the \emph{similar} sounds/objects. That
said, this difference in the level of similarity is accounted for in the
model through using the appropriate distance (Figure \ref{fig:stim}).

@@

\caption{Accuracy of word recognition as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both the models' predictions (dashed lines) and the experimental results (solid lines). The single-variance model uses one joint fitting parameter for both sound and meaning variances. The double-variance model uses two separate fittings parameters for the sound and the meaning variances. Error bars represent 95\% confidence intervals.}\label{fig:allData}
\end{figure}

Experimental results are shown in Figure \ref{fig:allData} (solid
lines). We first analyzed the results using a mixed-effects logistic
regression with sound distance, object distance and age group as fixed
effects, and with a maximal random effects structure (allowing us to
take into account the full nested structure of our data) (Barr, Levy,
Scheepers, \& Tily, 2013). We found main effects for all the fixed
effects in the regression. For the sound distance, we obtained
\(\beta =\) 0.52 (\(p\) \textless{} 0.001), replicating previous
findings that sound distance modulates success in word learning (e.g.,
Stager \& Werker, 1997). For object distance, we found \(\beta =\) 0.83
(\(p\) \textless{} 0.001), and this finding confirms the new prediction
of our model according to which object distance also modulates success
in word learning. Finally, for the age group, we obtained \(\beta =\)
0.76 (\(p\) \textless{} 0.001), showing that overall performance
improves with age.

\begin{table}

\caption{\label{tab:models}Characteristics and performance of the models used in this study.}
\centering
\begin{tabular}[t]{>{\bfseries}llrrrlrl}
\toprule
\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Children} & \multicolumn{2}{c}{Adults} \\
\cmidrule(l{2pt}r{2pt}){5-6} \cmidrule(l{2pt}r{2pt}){7-8}
Model & Structure & Param. & R\textsuperscript{2} & Sig\textsubscript{L} & Sig\textsubscript{C} & Sig\textsubscript{L} & Sig\textsubscript{C}\\
\midrule
model 1 & Sig\textsubscript{L} only & 1 & 0.27 & 1.00 & -- & 0.37 & --\\
model 2 & Sig\textsubscript{L} = Sig\textsubscript{C} & 1 & 0.95 & 0.60 & 0.6 & 0.15 & 0.15\\
model 3 & Sig\textsubscript{L} != Sig\textsubscript{C} & 2 & 1.00 & 0.83 & 0.31 & 0.12 & 0.17\\
\bottomrule
\end{tabular}
\end{table}

We next fit the three models obtained through expressions (1), (2), and
(3) to the participants' responses in each age group. The predictions of
the models are shown \ref{fig:allData} (dashed lines) and the parameter
estimates (for \(\sigma_L\) and \(\sigma_C\)) as well as models'
goodness to fit (i.e., measured through \(R^2\)) are presented in Table
\ref{tab:models}.

Model 1, which does not take into account ambiguity in recovering
concepts from observed objects, explains only a small part of the
variance. In contrast, Model 3, which does take into account this
ambiguity, accounts for all the variance. Interestingly, Model 2 which
has a single, shared uncertainty parameter for both auditory and visual
modalities still explains almost all the variance in human data,
suggesting that the explanatory power of the general Model 3 is largely
due to its structure, rather than its degrees of freedom.

As predicted, the uncertainty parameters were larger for children than
they were for adults (Table \ref{tab:models}), showing that the
probabilistic representations become more refined (that is, \(\sigma\)
becomes smaller) across development. Further, the parameter estimates of
Model 3 show that this developmental effect is more important for the
label-specific uncertainty that it is for the concept-specific
uncertainty.

@@

This paper explored the idea that some seemingly stage-like patterns in
cognitive development can be characterized in a continuous fashion. We
used as a case study the seminal work of Stager and Werker (1997)
showing a discrepancy between children's speech perception abilities and
their word learning skills. While this fact might seem like a specific
stage in early development, our model demonstrated, instead, that it can
be more simply understood in terms of continuous developmental change.

The main assumption of the model was that both word form and referent
are encoded in a probabilistic fashion. The model provided a
quantitative instantiation of the continuous development hypothesis
(Pajak et al., 2016; e.g., Swingley, 2007; Yoshida et al., 2009)). More
precisely, we showed that developmental changes in word-object mappings
can be characterized as a continuous refinement in the precision of the
probabilistic representations.

We find in the literature two broad accounts of development in the
Switch task: One that suggests \emph{direct} development of the sound
representation and one that hypothesizes \emph{indirect} development of
this representation through improvement in general cognitive resources.
On the first account, the sound representation becomes more precise as
learners refine the boundaries of their initially ambiguous phonetic
categories and as they gain more experience with the functional role of
these categories, i.e., contrasting word meaning (Apfelbaum \& McMurray,
2011; Dietrich, Swingley, \& Werker, 2007; Rost \& McMurray, 2009, 2010;
Yoshida et al., 2009). On the second account, the precision of sound
encoding in the switch task improves as a result of the maturation of
more general resources like the attentional and working memory capacity
(Hofer \& Levy, 2017; Stager \& Werker, 1997; Werker \& Fennell, 2004).
Such improvement allows older children and adults to better encode the
sound details while simultaneously matching these sounds to visual
objects. These two accounts are complementary and both seem to play a
role (see Tsui, Byers-Heinlein, \& Fennell, 2019 for a review).
